---
title: "Don't be K-Mean, join a gRoup"
author: "The Members of R Squad"
date: "10/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cluster)
library(factoextra)


# The following hook function is modified from https://bookdown.org/yihui/rmarkdown-cookbook/hook-truncate.html and it automatically truncates excessively long outputs for each code chunk by overwriting the default output method in the knitr package
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$out.lines)) {
    x <- xfun::split_lines(x)
    if (length(x) > n) {
      # truncate the output
      x <- c(head(x, n/2), "....\n", tail(x, n/2 + 1))
    }
    x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})
# Truncates outputs to 8 lines
knitr::opts_chunk$set(out.lines = 15) 
options(max.print = 50)
```

## Download Data

```{r data}
workshopQueryData = read.csv(url('https://drive.google.com/uc?id=1ff4xFh4fl0-SvpYNeYQoNvDbzdiZfn-t'))
bootcampTurnoutData = read.csv(url('https://drive.google.com/uc?id=1sovKLesEqPbkUte_ysRP9mGQ1gFCO2ME'))
workshopsData = read.csv(url('https://drive.google.com/uc?id=10MngpIZoAGgwAk_sxoORj7WPYs74nz5Y'))
applicantData = read.csv(url('https://drive.google.com/uc?id=1mPK8_AasPMdqy3D9D0kxjjKcIXmhhcQo'))
```

## Construct scoring matrix to cluster

```{r clstrMatrix}
cltrApplicantData = select(applicantData, c(userid, age_bin, classification, first_generation, 
                                            datascience_experience, technology_experience, num_hackathons_attended))
cltrApplicantData$age_bin = as.ordered(cltrApplicantData$age_bin)
cltrApplicantData$classification = ordered(cltrApplicantData$classification, levels = c('Fr', 'So', 'Jr', 'Sr', 'O', 'Ma', 'PhD'))
cltrApplicantData$first_generation = as.numeric(cltrApplicantData$first_generation)
for(i in 1:length(cltrApplicantData$technology_experience)) {
  cltrApplicantData$technology_experience[i]=length(strsplit(cltrApplicantData$technology_experience[i], ',')[[1]])
}
cltrApplicantData$num_hackathons_attended=ordered(cltrApplicantData$num_hackathons_attended, levels = c('0', '1-3', '4-7', '8-10', '10+'))
cltrApplicantData = data.frame(userid = cltrApplicantData$userid,data.matrix(select(cltrApplicantData, !userid)))
rownames(cltrApplicantData) = cltrApplicantData[,1]
cltrApplicantData = select(cltrApplicantData, !userid)
cltrApplicantData = scale(cltrApplicantData)
```

## Perform naive clustering with 4 clusters

```{r clustr4, message = FALSE, cache = TRUE}
distance = get_dist(cltrApplicantData)
fviz_dist(distance, gradient = list(low = 'red', mid = 'yellow', high = 'green'), show_labels = FALSE)
k4 = kmeans(cltrApplicantData, centers = 4, nstart = 25); k4
fviz_cluster(k4, data = cltrApplicantData, geom = 'point')
```

## Calculate the gap statistic to find the optimal number of clusters

```{r clustrGap, cache=TRUE, warning=FALSE, message = FALSE}
set.seed(123456789)
gapStat = clusGap(cltrApplicantData, FUN = kmeans, nstart = 25, K.max = 30, B = 50)
fviz_gap_stat(gapStat)
kOptimalGap = kmeans(cltrApplicantData, centers = 21, nstart = 25)
fviz_cluster(kOptimalGap, data = cltrApplicantData, geom = 'point')
```

## Sanity check: are people in the same cluster 'similar'?

```{r sanity}
clusteredApplicants = applicantData %>% mutate(cluster = kOptimalGap$cluster) %>% group_by(cluster)
#head(clusteredApplicants)
#table(clusteredApplicants$cluster)
filter(clusteredApplicants, cluster == 14) %>% select(!c(school, other_school, minors)) %>% sample_n(5)
filter(clusteredApplicants, cluster == 1) %>% select(!c(school, other_school, minors)) %>% sample_n(5)
```
Looking at cluster 14, it appears to be a group of people who are mostly upperclassmen or graduate level students with lots of experience and many technical skills.
In comparison, looking at cluster 1 it appears to be a group of people who are mostly underclassmen who are first gen students, very little experience and not many technical skills.

## Given a userid, recommend 5 group mates of similar ability

```{r groupmake, warning = FALSE, message = FALSE}
groupMake = function(user){
  userCluster = filter(clusteredApplicants, userid == user) %>% select(cluster)
  filter(clusteredApplicants, cluster == userCluster, userid != user) %>% select(userid) %>% sample_n(5) %>% return()
}
randomUserID = sample(applicantData$userid, 1)
groupMake(randomUserID)
```

